Mechanising Recursion Schemes with Magic-Free Coq Extraction

### 2024-10-10 paper/paper.tex
shrink related work slightly

\section{Related Work}
\label{sec:related-work}
%
%\begin{itemize}
%  \item McBride devised a generalisation of Bove and Capretta's general
%recursive function monad~\cite{McBride15, BoveC01}
%  \item Guarded recursion for productive definitions and general recursion
%~\cite{AtkeyM13, PaviottiMB15}
%  \item PACO library~\cite{HurNDV13}
%  \item Sized Types \cite{HughesPS96, abs-1202-3496}
%  \item Copatterns \cite{AbelPTS13, AbelP16}
%  \item Not that related: extraction for mutable arrays~\cite{SAKAGUCHI2020102372}
%        \item Extraction of Distributed Mobile code in Coq~\cite{MiculanP12}.
%\end{itemize}
Encoding recursion schemes in Coq is not new. We compare our work with other
encodings of program calculation techniques in Coq and recursion schemes in Coq
(Sections~\ref{sec:calccoq} and~\ref{sec:reccoq}). Additionally, we compare our
encoding of recursion schemes to other forms of termination checking for
nonstructural recursion (Section~\ref{sec:termcheck}), and we discuss other
approaches to fusion in functional programs (Section~\ref{sec:fusion}).

@@

\emph{Deforestation}, introduced by Wadler~\cite{wadler88}, is a specific
approach to fusion that aims to remove intermediate trees (such as lists)
created between chained operations. A well-known method of deforestation is
\haskell{foldr/build} fusion~\cite{gill93}, which optimizes list-processing by
fusing list-producing functions (like \haskell{map} or \haskell{filter}) with
list-consuming ones (like \haskell{foldr}), avoiding the creation of temporary
lists.

Extending this idea, stream fusion generalizes the concept to handle other data
types, using a lazy stream abstraction to represent sequences and fuse
transformations like \haskell{map} and \haskell{fold}~\cite{coutts07}. Both
approaches eliminate redundant structures and ensure computations happen in a
single efficient pass over the data.

Another dimension of fusion is \emph{tupling}, a technique used to combine
multiple recursive functions into a single function by grouping their results
into a tuple~\cite{hughes89}. This approach helps avoid redundant traversals of
the same data structure when multiple operations are applied. In addition to
techniques like \haskell{foldr/build} fusion and stream fusion, tupling
complements deforestation by allowing multiple recursive functions to share a
single traversal, further optimizing computations and minimizing intermediate
data structures.

>>>>

\section{Related Work}
\label{sec:related-work}
%
%\begin{itemize}
%  \item McBride devised a generalisation of Bove and Capretta's general
%recursive function monad~\cite{McBride15, BoveC01}
%  \item Guarded recursion for productive definitions and general recursion
%~\cite{AtkeyM13, PaviottiMB15}
%  \item PACO library~\cite{HurNDV13}
%  \item Sized Types \cite{HughesPS96, abs-1202-3496}
%  \item Copatterns \cite{AbelPTS13, AbelP16}
%  \item Not that related: extraction for mutable arrays~\cite{SAKAGUCHI2020102372}
%        \item Extraction of Distributed Mobile code in Coq~\cite{MiculanP12}.
%\end{itemize}
Encoding recursion schemes in Coq is not new. We compare our work with other
encodings of program calculation techniques and recursion schemes in Coq, and
discuss other approaches to fusion and to termination of nonstructural
recursion.
% and recursion schemes in Coq
% (Sections~\ref{sec:calccoq} and~\ref{sec:reccoq}). Additionally, we compare our
% encoding of recursion schemes to other forms of termination checking for
% nonstructural recursion (Section~\ref{sec:termcheck}), and we discuss other
% approaches to fusion in functional programs (Section~\ref{sec:fusion}).

@@

\emph{Deforestation}, introduced by Wadler~\cite{wadler88}, is a specific
approach to fusion that aims to remove intermediate trees (such as lists)
created between chained operations. A well-known method of deforestation is
\haskell{foldr/build} fusion~\cite{gill93}, which optimizes list-processing by
fusing list-producing functions (like \haskell{map} or \haskell{filter}) with
list-consuming ones (like \haskell{foldr}), avoiding the creation of temporary
lists.
%
Extending this idea, stream fusion generalizes the concept to handle other data
types, using a lazy stream abstraction to represent sequences and fuse
transformations like \haskell{map} and \haskell{fold}~\cite{coutts07}. Both
approaches eliminate redundant structures and ensure computations happen in a
single efficient pass over the data.

Another dimension of fusion is \emph{tupling}, a technique used to combine
multiple recursive functions into a single function by grouping their results
into a tuple~\cite{hughes89}. This approach helps avoid redundant traversals of
the same data structure when multiple operations are applied.
% In addition to
% techniques like \haskell{foldr/build} fusion and stream fusion, tupling
% complements deforestation by allowing multiple recursive functions to share a
% single traversal, further optimizing computations and minimizing intermediate
% data structures.
